Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-04-07 pkg/disttree/inst/draft/draft.Rnw
deleted not used color names
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2134 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 3641611a5da40f14767c78cddfcb88fbf2e7b8eb

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)
@

@@

<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
LM, GLM  
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
GAM 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
GAMLSS 
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\end{tikzpicture}
\caption{\label{fig:gam}Parametric modeling developments. (Generalized) linear models (left),
generalized additive models (middle), generalized additive models for location, scale, and shape (right).}
\end{figure}

@@

<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_distforest, echo=FALSE, results=hide>>=
## distforest
set.seed(4723)
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
xt <- 2*(x-0.5)
var <- exp(-(2*xt)^4)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
x <- c(1:nobs)/nobs
@

<<plot_motivation_distforest, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Regression tree 
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Random forest 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
Distributional forest 
\endminipage};
\node (g) at (0,-5.4) 
{
\minipage{0.29\textwidth}
\vspace{0.2cm}
%\hspace{0.0cm}
\centering
<<motivation_disttree, echo=FALSE, results=hide>>=
## disttree
set.seed(74)
x <- c(1:nobs)/nobs
var <- ytree/3
y <- ytrue + rnorm(nobs,0,var)
@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
lines(x = x, y = ytree, col = pal["forest"], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight["forest"], border = "transparent")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\vspace{-0.2cm}%\\
Distributional tree
\vspace{0.2cm}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{fig:tree}Tree and forest developments.
Regression tree (top left), distributional tree (bottom left), random forest
(top middle), and distributional forest (top right).}
\end{figure}

@@

\begin{figure}[t!]
\centering
<<rain_cross_axams_crps_skill_score, fig=TRUE, echo=FALSE, height=4.5, width = 7>>=
boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
        names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
        ylab = "CRPS skill score", col = "lightgray") 
abline(h = 0, col = pal["EMOS"], lwd = 2)
@
\caption{\label{fig:axams-crps}CRPS skill score from the 10 times 
7-fold cross validation at station Axams (1985--2012). The horizontal orange 
line pertains to the reference model EMOS.}
\end{figure}

@@

load("plot_map_rain/data/tirol.rda")
load("plot_map_rain/data/dem.rda")

data(StationsTyrol)
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(dem))
@

\begin{figure}[p!]
\centering
<<rain_all_crps_skill_score, fig=TRUE, echo=FALSE, height = 4.5, width = 7>>=
par(mar = c(3.1, 4.1, 1.1, 2.1))
matplot(t(s[,]), type = "l", lwd = 2, 
          col = gray(0.5, alpha = 0.2),
          lty = 1, axes = FALSE, 
          xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
lines(s[70,], col = pal[2], type = "o", pch = 19, lwd = 2)  
# Station Axams is the 77th station which is the 70th in the list of complete stations
  boxplot(s, add = TRUE, col = "transparent")
  abline(h = 0, col = pal["EMOS"], lwd = 2)

>>>>

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)
@

@@

<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x, ytrue, col = pal[1], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
LM, GLM  
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
GAM 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
GAMLSS 
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\end{tikzpicture}
\caption{\label{fig:gam}Parametric modeling developments. (Generalized) linear models (left),
generalized additive models (middle), generalized additive models for location, scale, and shape (right).}
\end{figure}

@@

<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal[1], lwd=7)
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_distforest, echo=FALSE, results=hide>>=
## distforest
set.seed(4723)
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
xt <- 2*(x-0.5)
var <- exp(-(2*xt)^4)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
x <- c(1:nobs)/nobs
@

<<plot_motivation_distforest, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Regression tree 
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Random forest 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
Distributional forest 
\endminipage};
\node (g) at (0,-5.4) 
{
\minipage{0.29\textwidth}
\vspace{0.2cm}
%\hspace{0.0cm}
\centering
<<motivation_disttree, echo=FALSE, results=hide>>=
## disttree
set.seed(74)
x <- c(1:nobs)/nobs
var <- ytree/3
y <- ytrue + rnorm(nobs,0,var)
@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
lines(x = x, y = ytree, col = pal[1], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight[1], border = "transparent")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\vspace{-0.2cm}%\\
Distributional tree
\vspace{0.2cm}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.9pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{fig:tree}Tree and forest developments.
Regression tree (top left), distributional tree (bottom left), random forest
(top middle), and distributional forest (top right).}
\end{figure}

@@

\begin{figure}[t!]
\centering
<<rain_cross_axams_crps_skill_score, fig=TRUE, echo=FALSE, height=4.5, width = 7>>=
par(mar = c(2.5,2,1,2))
boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
        names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
        ylab = "CRPS skill score", col = "lightgray") 
abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:axams-crps}CRPS skill score from the 10 times 
7-fold cross validation at station Axams (1985--2012). The horizontal orange 
line pertains to the reference model EMOS.}
\end{figure}

@@

#load("plot_map_rain/data/tirol.rda")
#load("plot_map_rain/data/dem.rda")

data(StationsTyrol, package = "RainTyrol")
data(MapTyrol, package = "RainTyrol")
data(MapTyrol_border, package = "RainTyrol")
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(dem))
@

\begin{figure}[p!]
\centering
<<rain_all_crps_skill_score, fig=TRUE, echo=FALSE, height = 4.5, width = 7>>=
par(mar = c(3.1, 4.1, 1.1, 2.1))
matplot(t(s[,]), type = "l", lwd = 2, 
          col = gray(0.5, alpha = 0.2),
          lty = 1, axes = FALSE, 
          xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
lines(s[70,], col = pal[2], type = "o", pch = 19, lwd = 2)  
# Station Axams is the 77th station which is the 70th in the list of complete stations
  boxplot(s, add = TRUE, col = "transparent")
  abline(h = 0, col = pal[5], lwd = 2)
